---
title: "Project2 Practice"
author: "Muriel Pokol"
date: "October 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup2, include=FALSE}
require(rgdal)
library(ggplot2)
require(leaflet)
require(dplyr)
require(readxl)
library(jsonlite) # fromJSON
library(utils) # URLencode functions
require(stringr)
require(leaflet.extras)

```

```{r, data_load, results='hide'}
railway_data     <- readOGR("./GBR_rails.shp", GDAL1_integer64_policy=TRUE)  # for lines map
admin_data_excel <- read.csv("GBR_adm2.csv")                                 # for poly map
admin.load       <- readOGR("./GBR_adm2.shp",  GDAL1_integer64_policy=TRUE)  # for poly map  
```

## YAY APIs
Data from : https://data.gov.uk/dataset/e284af14-70f0-43b4-a49f-6010eda83fed/travel-to-work-areas-december-2011-ultra-generalised-clipped-boundaries-in-united-kingdom
```{r, esri}
# Travel from Work Areas
# Load as geojson
travel <- readOGR("http://geoportal1-ons.opendata.arcgis.com/datasets/d3062ec5f03b49a7be631d71586cac8c_4.geojson")
# Plot polling locations
plot(travel)
```

## Train Lines across the United Kingdom
Builds a map with layer of lines.  This map also includes a functioning `layersControl()` to choose between two basemaps. [^1]
 
[^1]: Railway data did not download with a csv attachment denoting *which* train lines they are or if the railways are named, and so there is no popup associated with this dataset. I spoke to Malvika about this in Office Hours. 

```{r, lines}
leaflet() %>%
  
  # Determine base maps
  addProviderTiles("Stamen.Toner",      group = "Toner Map") %>%
  addProviderTiles("Esri.WorldTerrain", group = "World Terrain Map") %>%
  addProviderTiles("OpenStreetMap.HOT", group = "Street Map") %>% 
  
  addPolylines(data = railway_data, color = "tomato", highlightOptions = highlightOptions(color="White", 
                                                                     bringToFront=TRUE)) %>%
  addPolygons(data = travel, color = "blue", fillOpacity = 0.25, 
              highlightOptions = highlightOptions(color="White", bringToFront=TRUE)) %>%
  
  # Add in layers control
  addLayersControl(
    baseGroups = c("Toner Map", "World Terrain Map", "Street Map"),
    options = layersControlOptions(collapsed=FALSE)
  )
```

<br>

## Administrative Counties in the United Kingdom 
Builds a map with layer of polygons, again with a functioning `layersControl()` to choose between two basemaps. Counties are broken down with the polygon boundaries, and county name is shown as a popup.
```{r, merge}
# start csv ID count at 0 instead of 1 to match shp IDs
admin_data_excel$OBJECTID = admin_data_excel$OBJECTID - 1

# Merge data on thier respective IDs
admin      <- admin.load[admin.load$ID_2 %in% admin_data_excel$OBJECTID,]
admin@data <- merge(admin@data, admin_data_excel, sort=FALSE,
                    by.x='ID_2', by.y='OBJECTID')
```

```{r, polygons}
leaflet(data = admin) %>% 
  
  # Determine base maps
  addProviderTiles("Stamen.Toner",    group = "Toner Map") %>%
  addProviderTiles("Esri.WorldTerrain", group = "World Terrain Map") %>%
  addProviderTiles("OpenStreetMap.HOT", group = "Street Map") %>%
  
  addPolygons(color = "tomato", fillOpacity = 0.5, 
              highlightOptions = highlightOptions(color="White", bringToFront=TRUE),
              popup = ~paste0(admin$NAME_2.x, ", ", admin$TYPE_2.x)) %>%
  addPolylines(data = travel, color = "blue", fillOpacity = 0.25, 
              highlightOptions = highlightOptions(color="White", bringToFront=TRUE)) %>%
  
  # Add in layers control
  addLayersControl(
    baseGroups = c("Toner Map", "World Terrain Map", "Street Map"),
    options = layersControlOptions(collapsed=FALSE)
  )
```



```{r}
ggplot(admin_data_excel, aes(x=NAME_1, fill=TYPE_2)) + geom_bar()
```





